---
title: "Using the Binomial Distribution in R"
# author: "John Minter"
date: "Last updated: 2018-06-22"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Using the Binomial Distribution in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
```
# An example application

This example is adapted from
[www.r-tutor.com](http://www.r-tutor.com/elementary-statistics/probability-distributions/binomial-distribution).

The binomial distribution is a discrete probability distribution. It
describes the outcome of n independent trials in an experiment. Each
trial is assumed to have only two outcomes, either success or failure.
If the probability of a successful trial is `p`, then the probability
of having x successful outcomes in an experiment of n independent trials
given by.

$$f(x) = \dbinom{n}{x} \ p^x (1-p)^{(n-x)} \ where \ x=1,2,\ldots,n$$

## Problem

Suppose there are twelve multiple choice questions in an English class
quiz. Each question has five possible answers, and **only one** of them
is correct. Find the probability of having four or less correct answers
if a student attempts to answer every question at random.

## Solution

Since only one out of five possible answers is correct, the probability
of answering a question correctly by random is $1/5 = 0.2$. We can use R
to find the probability of having exactly 4 correct answers by random
attempts as follows.

```{r 4CorrectAns}
ans <- dbinom(4, size=12, prob=0.2)
print(ans)
```

To find the probability of having **four or less correct answers**
by random attempts, we can **sequentially apply** the function `dbinom`
with $x=0,1,\ldots,4$.

```{r, BruteForce}

ans <-  dbinom(0, size=12, prob=0.2) + 
        dbinom(1, size=12, prob=0.2) + 
        dbinom(2, size=12, prob=0.2) + 
        dbinom(3, size=12, prob=0.2) + 
        dbinom(4, size=12, prob=0.2)
print(ans)
```

Alternatively, we can use the **cumulative probability function for
binomial distribution**, `pbinom`.

```{r usePbinom}

ans <- pbinom(4, size=12, prob=0.2)
print(ans)
```

## Answer

The probability of **four or less** questions answered correctly by
random in a twelve question multiple choice quiz is
`r sprintf("%.1f", round(100*ans, 1))`%.

# Plotting the binomial distribution

This example is adapted from
[stackoverflow](https://stackoverflow.com/questions/14933884/statistics-binomial-distribution-and-simple-graph-in-r).

## The problem

What is the probability of having **at least 16 successes** out of 20
operations **given that the probability of success is 0.8**?

## The solution

We will use `ggplot2` to create the basic plot

```{r plotBasicBinomial, fig.width=7.125}
require(ggplot2)

df <- data.frame(x=1:20, prob=dbinom(1:20, 20, prob=0.8))
plt <-  ggplot(data = df, aes(x = x, y = prob)) + geom_line() +
        geom_ribbon(data = subset(df, x >= 16 & x <= 20),
        aes(ymax = prob), ymin=0, fill = "red",
            colour = NA, alpha = 0.5)
print(plt)
```

We can make the plot more readable with better axis labels.

```{r plotBetter, fig.width=7.125}
plt  <-  plt +
         xlab("Number of successful operations") +
         ylab("Probability") +
         ggtitle("Probability of 16/20 successful operations") +
         theme(axis.text=element_text(size=12),
               axis.title=element_text(size=12),
               plot.title=element_text(hjust = 0.5))
print(plt)
```

