---
title: "Using the Binomial Distribution in R"
# author: "John Minter"
date: "Last updated: 2018-06-03"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the Binomial Distribution in R}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
# suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
```
# Introduction

This example is adapted from
[www.r-tutor.com](http://www.r-tutor.com/elementary-statistics/probability-distributions/binomial-distribution).

The binomial distribution is a discrete probability distribution. It
describes the outcome of n independent trials in an experiment. Each
trial is assumed to have only two outcomes, either success or failure.
If the probability of a successful trial is `p`, then the probability
of having x successful outcomes in an experiment of n independent trials
given by.

$$f(x) = 	\dbinom{n}{x} \ p^x (1-p)^{(n-x)} \ where \ x=1,2,\ldots,n$$

# Problem

Suppose there are twelve multiple choice questions in an English class
quiz. Each question has five possible answers, and **only one** of them
is correct. Find the probability of having four or less correct answers
if a student attempts to answer every question at random.

# Solution

Since only one out of five possible answers is correct, the probability
of answering a question correctly by random is $1/5 = 0.2$. We can use R
to find the probability of having exactly 4 correct answers by random
attempts as follows.

```{r 4CorrectAns}
ans <- dbinom(4, size=12, prob=0.2)
print(ans)
```

To find the probability of having **four or less correct answers**
by random attempts, we can **sequentially apply** the function `dbinom`
with $x=0,1,\ldots,4$.

```{r, BruteForce}

ans <- dbinom(0, size=12, prob=0.2) + 
       dbinom(1, size=12, prob=0.2) + 
       dbinom(2, size=12, prob=0.2) + 
       dbinom(3, size=12, prob=0.2) + 
       dbinom(4, size=12, prob=0.2)
print(ans)
```

Alternatively, we can use the **cumulative probability function for
binomial distribution**, `pbinom`.

```{r usePbinom}

ans <- pbinom(4, size=12, prob=0.2)
print(ans)

```

# Answer

The probability of **four or less** questions answered correctly by
random in a twelve question multiple choice quiz is
`r sprintf("%.1f", round(100*ans, 1))`%.

