---
title: "Using a boxplot to visualize data and discover outliers"
# author: "John Minter"
date: "Last updated: 2018-06-07"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Using a boxplot visualize data and discover outliers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
```
# Introduction

Box plots are useful for visualizing data

# Example 1 - A single vector of data

## Introduction

This example is adapted from data supplied by
[https://www.statcan.gc.ca](https://www.statcan.gc.ca/edu/power-pouvoir/ch8/5214816-eng.htm#a3).

## Analysis

### Input the data

The masses (to the nearest tenth of a kilogram) of 30 students were measured
and recorded as follows:

```{r inputData}
df <- data.frame(mass = c(59.2, 61.5, 62.3, 61.4, 60.9, 59.8,
						  60.5, 59.0, 61.1, 60.7, 61.6, 56.3,
						  61.9, 65.7, 60.4, 58.9, 59.0, 61.2,
						  62.1, 61.4, 58.4, 60.8, 60.2, 62.7,
						  60.0, 59.3, 61.9, 61.7, 58.4, 62.2))
```

### Make the boxplot 

**Note:** We only have one list of data, so when we use `ggplot2`
we set the `x` axis to `""`.

```{r makePlot, fig.width=6.5}
library(ggplot2)


p <- ggplot(df) + aes(x="", y=mass) +
	 geom_boxplot() +
	 ggtitle("Body mass of 30 students") +
	 labs(x="", y="Body mass [kg]") +
	 theme(plot.title = element_text(size = 14,
									 family = "Tahoma",
									 face = "bold",
									 hjust = 0.5),
		 	panel.border = element_rect(colour = "black",
									   fill=NA, size=.5),
			text = element_text(size = 12, family = "Tahoma"),
			axis.title = element_text(face="bold"),
			axis.text.x=element_text(size = 11))

print(p)
```

### Summary

Note the outlier just under 66 kg. Also note that the whiskers have
different lengths.

# Example 2 - A more complicated example with multiple factors.

This is adapted from the `ggplot2` examples
[here](http://ggplot2.tidyverse.org/reference/geom_boxplot.html),

## Choose a data set

We will use the standard `mpg` data set that comes with R.

```{r summarizeData}
summary(mpg)

```

## Make the plot

```{r ggplotMPG, fig.width=6.5}
plt <-	ggplot(mpg, aes(class, hwy)) +
		geom_boxplot() +
		ggtitle("Highway mpg of cars") +
		labs(x="car class", y="Highway mpg") +
		theme(plot.title = element_text(size = 14,
										family = "Tahoma",
										face = "bold",
										hjust = 0.5), # center title
			  panel.border = element_rect(colour = "black",
										  fill=NA, size=.5),
			  text = element_text(size = 12, family = "Tahoma"),
			  axis.title = element_text(face="bold"),
			  axis.text.x=element_text(size = 11))

print(plt)

```

## now let's superimpose the data

We will extend the plot. Note the use of the `jitter` argument to spread
out the data along the x-axis.

```{r withJitter, fig.width=6.5}
plt <-	plt +
		geom_boxplot() +
		geom_jitter(width = 0.2)
print(plt)
```
