---
title: "Confidence Intervals"
# author: "John Minter"
date: "Last updated: 2018-06-08"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Confidence Intervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(mosaic))))
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(plotly))))
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(openintro))))
library(png)
library(grid)
library(statshelpR)
data(meanTime25)
```
# Introduction

A point estimate provides a single plausible value for a parameter.
However, a point estimate is rarely perfect. Usually there is some
error in the estimate. To account for this error, we provide a
plausable range of values for the parameter.

# An approximate 95% confidence level

A single point estimate such as the **sample mean** will not likely be
equal to the exact **population parameter** like the true population
mean. A plausable range of values is called a **confidence interval**.

## The standard error

The **standard error** is a measure of the uncertainty associated with
the point estimate and provides a guide for how large we should make the
confidence level. The standard error represents the standard deviation
associated with the estimate.

If the interval spreads out 2 standard errors from the point estimate,
we can be roughly 95% **confident** that we have captured the true
parameter.

$$point\ etimate \pm 2 \times SE$$

What does 95% confident mean? Suppose we took many samples and built a
95% confidence interval from each sample. Then 95% of those intervals
would contain the actual mean.

```{r showImg,fig.width=7.5,echo=FALSE}
manySampImg <- readPNG('inc/many-samples.png')
grid.raster(manySampImg)
```
Note that **only one of these 25 estimates**
**did not capture the true mean of the population**. 96% of the
samples captured the true mean.

```{r SimNormal, fig.width=7.5}
df <- data.frame(obs=meanTime25)


plt <-	ggplot(data=df, aes(obs)) + 
		geom_histogram(breaks=seq(85, 105, by = .5), col="black",
					   fill="blue", alpha = .5) +
		ylab(label="frequency") + xlab("sample mean") +
		ggtitle("Sample Mean Distribution") +
		theme(axis.text=element_text(size=12),
			  axis.title=element_text(size=14),
			  plot.title = element_text(hjust = 0.5)) # center the title

print(plt)

```

And we can compute a **normal probability plot**:

```{r qqplotl, fig.width=7.5}
qqPlt <- ggplot(data=df, aes(sample=obs))+ stat_qq() +
		 ylab(label="sample means") + xlab("theoretical quantiles") +
		 ggtitle("Normal Probability Plot") +
		 theme(axis.text=element_text(size=12),
		 	   axis.title=element_text(size=14),
			   plot.title = element_text(hjust = 0.5)) # center the title
		

print(qqPlt)
```



# A sampling distribution for the mean

# Changing the confidence level

# Proper interpretation of confidence levels


