---
title: "Computing the Confidence Level to use for $t^*$"
# author: "John Minter"
date: "Last updated: 2018-06-07"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Computing the Confidence Level to use for T-star}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
```

# Introduction

The [stats4stem.org](http://www.stats4stem.org/r-t-distribution)
website provides a table with the desired confidence level for
a $t^*$ calculation and the the confidence level entered in R's
`qt()` function. This is presented without explanation.

This is an exploratory data analysis

We can load the data in R as shown below.

```{r loadData}
library(statshelpR)
library(pander)

data(tstar_cl_r)
pander(tstar_cl_r)
```

We can plot the data and fit it to a linear model

```{r plotData, fig.width=6.5}

plt <-	ggplot(tstar_cl_r, aes(x=CL, y=R)) +
		geom_point(color='black', size=2) +
		xlab("CL") + ylab("R") +
		ggtitle("R as a function of CL") +
		theme(axis.text=element_text(size=12),
			  axis.title=element_text(size=14),
			  plot.title = element_text(hjust = 0.5))

print(plt)
```

Let's do the linear model

```{r linModel}
fit <- lm(R ~ CL, data = tstar_cl_r)
pander(summary(fit))
```

And let's update our plot with the predicted line.

```{r updatePlot, fig.width=6.5}

pred <- data.frame(R = predict(fit, tstar_cl_r))
pred$CL <- tstar_cl_r$CL
pander(pred)

plt <- plt +  geom_line(color='red',data = pred, aes(x=CL, y=R))

print(plt)
```

So our equation is $R = 0.5*CL + 0.5$

```{r testSoln}
CL <- tstar_cl_r$CL

CL_to_R <- function(x){
	R <- 0.5 * x + 0.5
	return(R)
}


rVal <- sapply(CL, CL_to_R)

dfCalc <- data.frame(CL=CL, R=rVal)
pander(dfCalc)

```
