---
title: "Confidence Intervals"
# author: "John Minter"
date: "Last updated: 2018-06-22"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Confidence Intervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(mosaic))))
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(plotly))))
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(openintro))))
library(png)
library(grid)
library(statshelpR)
library(pander)
data(meanTime25)
```
# Introduction

This is adapted from YouTube video from 
[OpenIntro.Org](https://www.youtube.com/watch?v=FUaXoKdCre4) and some
information from [davidmlane.com](http://davidmlane.com/hyperstat/A103735.html).

A point estimate provides a single plausible value for a parameter.
However, a point estimate is rarely perfect. Usually there is some
error in the estimate. To account for this error, we provide a
plausible range of values for the parameter.

# An approximate 95% confidence level

A single point estimate such as the **sample mean** will not likely be
equal to the exact **population parameter** like the true population
mean. A plausible range of values is called a **confidence interval**.

## The standard error

The **standard error** is a measure of the uncertainty associated with
the point estimate and provides a guide for how large we should make the
confidence level. The standard error represents the standard deviation
associated with the estimate.

The standard error of the mean is designated as: $\sigma_M$. It is the
standard deviation of the sampling distribution of the mean. The formula
for the standard error of the mean is:

$$\sigma_M = \frac{\sigma}{\sqrt{N}}$$

where $\sigma$ is the standard deviation of the original distribution
and N is the sample size (the number of scores each mean is based upon).
This formula **does not assume a normal distribution**. *However*, many
of the uses of the formula **do assume a normal distribution**.

The formula shows that **the larger the sample size*, the
**smaller the standard error of the mean**. More specifically,
the size of the standard error of the mean is **inversely proportional**
to the **square root of the sample size**.

A plot of the effect of sample size on the standard error the standard
error for a standard deviation of 10 is show below.

```{r stdErrorSampleSize, fig.width=7.5}
sigma <- 10
n <- seq(1, 30, 1)
func <- function(n){
  return(sigma/sqrt(n))
}

se <- sapply(n,func)
df <- data.frame(n=n, se=se)

plt <-  ggplot(data=df, aes(x=n,y=se)) + geom_point() +
        xlab(label="number of samples") + ylab("SE") +
        scale_y_continuous(breaks = seq(from = 0, to = 10, by = 1),
                      	   limits = c(0,10)) +
        ggtitle(expression(paste(sigma[M], " as a function of sample size"))) +
        theme(axis.text=element_text(size=12),
              axis.title=element_text(size=14),
              plot.title = element_text(hjust = 0.5)) # center the title
print(plt)
```

If the interval spreads out 2 standard errors from the point estimate,
we can be roughly 95% **confident** that we have captured the true
parameter.

$$point\ estimate \pm 2 \times SE$$

# A sampling distribution for the mean

What does 95% confidence mean? Suppose we took many samples and built a
95% confidence interval from each sample. Then 95% of those intervals
would contain the actual mean.


```{r showImg,fig.width=7.5,echo=FALSE}
manySampImg <- readPNG('inc/many-samples.png')
grid.raster(manySampImg)
```
Note that **only one of these 25 estimates**
**did not capture the true mean of the population**. 96% of the
samples captured the true mean.

Let's take 100,000 samples, calculate the mean of each, and plot
them in a histogram to get an especially accurate depiction of the
sampling distribution.

```{r SimNormal, fig.width=7.5}
df <- data.frame(obs=meanTime25)

plt <- ggplot(data=df, aes(obs)) + 
       geom_histogram(breaks=seq(85, 105, by = .5), col="black",
                      fill="blue", alpha = .5) +
        ylab(label="frequency") + xlab("sample mean") +
        ggtitle("Sample Mean Distribution") +
        theme(axis.text=element_text(size=12),
              axis.title=element_text(size=14),
              plot.title = element_text(hjust = 0.5)) # center the title
print(plt)
```

And we can compute a **normal probability plot** to show that the
distribution is indeed normal.

```{r qqplotl, fig.width=7.5}
qqPlt <- ggplot(data=df, aes(sample=obs))+ stat_qq() +
         ylab(label="sample means") + xlab("theoretical quantiles") +
         ggtitle("Normal Probability Plot") +
         theme(axis.text=element_text(size=12),
               axis.title=element_text(size=14),
               plot.title = element_text(hjust = 0.5)) # center the title
print(qqPlt)
```

Both plots show that the collection of means follow a normal model.
This result can be explained by the **Central Limit Theorem**.

An **informal description** is

> If a sample consists of **at least 30 independent observations**
> and the **data are not strongly skewed**, then the sample mean is
> well approximated by a normal model.

# Changing the confidence level

Perhaps we want a 99% confidence level, we would need to widen our
interval. If we want a lower confidence level, we could narrow our
interval.

There are 3 components:

1. **The point estimate**: The 95% confidence interval structure
provides guidance in how to make intervals with new confidence
levels: point estimate $\pm 1.96 \times SE$ To create a 99%
confidence interval we change the point estimate to 
$\pm 2.58 \times SE$. Our `statshelpR` package supplies the
function `calc_z_star` to compute this value. An example to
compute $Z^*$ for a 99% confidence interval is shown below:

```{r calc_z_star_99_pct}

z_star <- calc_z_star(0.99)
names(z_star) <- "z_star"
pander(z_star)
```

2. **Conditions required** for the sampling distribution of $\bar{x}$
is nearly normal and the estimate of the standard error is sufficiently
accurate.

- The sample observations are **independent**. These include random
assignment in an experiment or that they are obtained by a random sample
of < 10% of the population.

- The sample size is large. $n > 30$ is a good rule of thumb.

- The distribution of observations is not **strongly skewed.**

# Proper interpretation of confidence levels

We need to be careful with our language in expressing confidence
levels. An example is "We are x% confident that the population
parameter is between the lower and upper bound of our confidence
interval."

- *Incorrect* language might try to describe the confidence interval
as capturing the population with a certain probability. This is one
of the most common errors. The probability is that the parameter is only
within our boundaries.

- Confidence intervals *only try to capture the population parameter.*
A confidence interval **says nothing** about the confidence of capturing 
**individual observations**, a **proportion of observations**, or about
capturing **point estimates**.
