---
title: "Making a dotplot"
# author: "John Minter"
date: "Last updated: 2018-06-22"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Making a dotplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
```
# Introduction

Dotplots were made popular by the statistician W. S. Cleveland and are
described in the book [The Elements of Graphing Data](https://www.amazon.com/Elements-Graphing-Data-William-Cleveland/dp/0963488414/).

The R core team implemented this function in the base graphics package as
the function `stripchart`. You can get help on this function by typing
the command `??stripchart`.

There is a helpful YouTube video [here](https://www.youtube.com/watch?v=x9_rrjVxGdY).

# Example

## Introduction

This example is used data supplied from page 41 in your text
[Robert Johnson and Patricia Kuby](https://www.amazon.com/Elementary-Statistics-Available-Titles-Aplia/dp/0538733500/).

## Analysis

### Input the data

Use the list of 19 exam grades from **Illustration 2.3** on p. 41.
We will create a list named `grades`.

```{r inputData}
grades <- c(76, 74, 82, 96, 66, 76, 78, 72, 52, 68,
            86, 84, 62, 76, 78, 92, 82, 74, 88)
```

### Make the dotplot

First, use the base graphics function `stripchart`:

```{r stripChart,fig.width=6.5}
stripchart(grades, method="stack", at=c(0.05), pch=20, cex=3, las-1,
           frame.plot=FALSE, xlim=c(50,100), main="Exam grades")
```

We can also try the newer `ggplot2` graphics.

```{r ggplot2DP, fig.width=6.5}

df <- data.frame(grades=grades)
a <-  ggplot(df, aes(x = grades)) +
      geom_dotplot(binwidth=1) +
      scale_x_continuous(breaks=seq(from=50, to=100, by=5),
                         limits = c(50, 100)) +
      scale_y_continuous(NULL, breaks = NULL) +
      xlab("grade") +
      theme(axis.text=element_text(size=12),
            axis.title=element_text(size=12),
            plot.title=element_text(hjust = 0.5))
print(a)
```



For comparison, let's look at the histogram using the base graphics
function. We will set the breaks to have bins one grade wide. In
general, histograms are more useful, especially for larger data sets.

```{r makeHisto, fig.width=6.5}
minG <- min(grades)-1
maxG <- max(grades)+1
breaks <- seq(minG, maxG, 1)
a <- hist(grades, breaks, xlim=c(minG, 100))
```

Now try the newer ggplot2 approach

```{r makeGGHisto, fig.width=6.5}
a <- ggplot(df, aes(x = grades)) +
     geom_histogram(bins=50) +
     xlab("grade") + 
     ggtitle("Exam grade distribution") +
     theme(axis.text=element_text(size=12),
           axis.title=element_text(size=12),
           plot.title=element_text(hjust = 0.5)) # center the title
print(a)
```

