---
title: "Hypothesis testing of the variance"
author: "John Minter"
date: "Last updated: 2018-05-19"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hypothesis testing of the variance}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
```
# Introduction

A vendor claims that the variation of a batch of washers is within
$\pm 0.002$ cm. The customer measured the thickness of 15 randomly
selected washers and reported the data (in cm) below.

```{r inputData}
dat <- c(0.422, 0.424, 0.424, 0.430, 0.419,
         0.424, 0.420, 0.424, 0.425, 0.425,
         0.423, 0.431, 0.437, 0.422, 0.434)
```
Does this sample of washers meet the variance criterion at the 
95% confidence level? We will compute the 95% confidence interval
of our sample.

```{r computeVar"}
library(statshelpR)
library(knitr)

df <- var.interval(dat, 0.002, 0.95)
kable(df)
```

# Summary

Note that the 95% confidence interval for the standard deviation
`r df$Value[7]` - `r df$Value[6]` and does not contain 0.002. Therefore
the batch us unacceptable.
