---
title: "Linear regression analysis of height data"
# author: "John Minter"
date: "Last updated: 2018-06-08"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Linear regression analysis of height data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(mosaic))))
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(plotly))))
```
# Introduction

Linear regression is a useful technique in many sciences. One of the
first studies was performed by Francis Galton, the nephew of
Charles Darwin. Galton wanted to study the relationship between the
height of parents and their children.

We can get the data from the `mosaic` package.

```{r getData, warning=FALSE}
library(mosaic)
library(pander)
data(Galton)
pander(head(Galton, 10))

```
We want to add a column of data for the `average parent height`

```{r}
Galton$midparentheight=(Galton$father+Galton$mother)/2
```

Let's plot the data and let `ggplot2()` add the fits so we can
easily see the results. We will use the `facet_wrap()` option
to separate male and female genders.


```{r plotResults, fig.width=7.25, warning=FALSE, message=FALSE}
library(grid)
plt <-	ggplot(data = Galton, aes(x = height, y = midparentheight)) +
		geom_point(aes(colour = sex), size = 1.5) +
        geom_smooth(aes(colour = sex, fill = sex), method = "lm" ,size = .5) +
		facet_wrap(~ sex) +
		ggtitle("Galton height data with linear models") +
		theme(panel.spacing = unit(2, "lines"),
			  axis.text=element_text(size=12),
			  axis.title=element_text(size=12),
			  plot.title=element_text(hjust = 0.5)) # center the title

print(plt)
```

Let's look more closely at the fits.

```{r testFits}
female <- Galton[which(Galton$sex=='F'), ]
femaleLM <- lm(height ~ midparentheight, data = female)
pander(summary(femaleLM))

male <- Galton[which(Galton$sex=='M'), ]
maleLM <- lm(height ~ midparentheight, data = male)
pander(summary(maleLM))

```

We see that both fits are statistically significant for their
**respective populations**. That said, the **variance** in the
population tells us that the model cannot tell us much about
a **single individual**.
