---
title: "Probability: The general additivity rule"
# author: "John Minter"
date: "Last updated: 2018-06-06"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Probability: The general additivity rule}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
```
# Introduction to probability

This example is adapted from an example in the first 9:30 min from the
OpenIntro
[YouTube](https://youtu.be/rG-SLQ2uF8U?list=PLkIselvEzpM5EgoOajhw83Ax_FktnlD6n)
video.

Start with a fair die. It has six sides. What is the probability of
rolling a 1?     
$P(1) = \frac{1}{6}$. Indeed the probability of all the
other numbers is $P(n) = \frac{1}{6}$.

We can also ask, what is the probability of rolling a 1 or a 2?
$P(1 \ or\ 2) = \frac{2}{6} = \frac{1}{3}$ since all of the numbers
are equally probable.

Probability is used to answer the question of how likely something is
to happen. Usually it is used to model **random processes**. It can also
be used for **non-random processes**. The stock market is one such
example. It looks random but each event is the result of someone's choice
about what they want to buy or sell. But the system as a whole looks random.

# Disjoint or mutually-exclusive event

First we consider **disjoint** or **mutually-exclusive** events.
Rolling a single die is one example. We only get 1 result for a
single roll. There is a rule associated with disjoint events,
like rolling a die. If we want to know:    

$P(1\ or\ 2) = P(1) + P(2) = \frac{1}{3}$

because each of the outcomes has the same probability. We could also
extend this for more mutually exclusive outcomes such as    

$P(1\ or\ 2\ or\ 3\ or\ 5) = P(1) + P(2) + P(3) + P(5) = \frac{4}{6} = \frac{2}{3}$.

We are using what is called the **Addition Rule for Disjoint Events**.

Next we want to knpw what happens with events that are **not Disjoint**.




# Dealing with events that are not disjoint

## Using a standard card deck

We want to consider an example where we draw cards from a standard
card deck, shown below.

```{r cardDeck, fig.width=7, fig.height=1.25, fig.cap="A standard 52 card deck containing clubs, diamonds, hearts, and spades", echo=FALSE}
library(png)
library(grid)
img <- readPNG("./inc/std-card-deck.png")
grid.raster(img)
```

## Problem Statement

Compute the probability of drawing a **diamond** or a **face card**

## A Venn Diagram helps

```{r VennDiagram, fig.align='center', fig.width=4, fig.cap="A Venn Diagram of the problem",echo=FALSE}

img <- readPNG("./inc/card-example-venn-diagram.png")
grid.raster(img)
```

Note the following probabilities:

$P(diamond) = P(D) = \frac{13}{54} = \frac{1}{4} = 0.25$    
$P(face\ card) = P(FC) = \frac{12}{52} = \frac{3}{13} \approx 0.231$

If we want to know $P(diamond\ or\ face\ card)$ we
**can't use the additivity rule for disjoint events** because
**not all face cards are diamonds**.

## The General Additivity Rule

$P(A \ or\ B) = P(A) + P(B) - P(A\ and \ B)$

so our answer is

$P(D \ or\ FC) = \frac{13}{52} + \frac{12}{52} - \frac{3}{52} = \frac{22}{52} = \frac{11}{26} \approx 0.423$

and is consistent with our Venn Diagram above.

