---
title: "Probability: The general additivity rule"
# author: "John Minter"
date: "Last updated: 2018-06-06"
output:
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Probability: The general additivity rule}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
suppressWarnings(suppressMessages(suppressPackageStartupMessages(library(ggplot2))))
```
# Introduction

This example is adapted from an example from the OpenIntro
[YouTube](https://youtu.be/rG-SLQ2uF8U?list=PLkIselvEzpM5EgoOajhw83Ax_FktnlD6n)
channel.

# The Problem

## Use a standard card deck

We want to consider an example where we draw cards from a standard
card deck, shown below.

```{r cardDeck, fig.width=7, fig.height=1.25, fig.cap="A standard 52 card deck containing clubs, diamonds, hearts, and spades", echo=FALSE}
library(png)
library(grid)
img <- readPNG("./inc/std-card-deck.png")
grid.raster(img)
```

## Problem Statement

Compute the probability of drawing a **diamond** or a **face card**

## A Venn Diagram helps

```{r VennDiagram, fig.align='center', fig.width=4, fig.cap="A Venn Diagram of the problem",echo=FALSE}

img <- readPNG("./inc/card-example-venn-diagram.png")
grid.raster(img)
```

Note the following probabilities:

$P(diamond) = P(D) = \frac{13}{54} = \frac{1}{4} = 0.25$    
$P(face\ card) = P(FC) = \frac{12}{52} = \frac{3}{13} \approx 0.231$    

## The General Additivity Rule

$P(A \ or\ B) = P(A) + P(B) - P(A\ and \ B)$

so our answer is

$P(D \ or\ FC) = \frac{13}{52} + \frac{12}{52} - \frac{3}{52} = \frac{22}{52} = \frac{11}{26} \approx 0.423$

and is consistent with our Venn Diagram above.

